---
title: "Show_NP"
author: "CJR"
date: "5/18/2022"
output: html_document
---

##Setup and Data Trim
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(gratia)
library(mgcv)
library(tidymv)

```

```{r data ingest}


det_raw <- read_csv("C:\\Users\\robbi\\Dropbox\\Detrital Nutrients Synth\\DetNutSynth_Database.csv")

```



```{r create new summary variables, include=FALSE}
det <- det_raw%>%
  group_by(First_Author, Publication_Title, Time_Series_ID)%>%
  mutate(series_index = group_indices())%>%#creates a unique index for each time series
  ungroup()%>%
  group_by(series_index)%>% #functionally, used to confine any window functions like first() to a time series rather than the first observation of whole dataframe
  mutate(initial_C = first(C_per),
         initial_N = first(N_per),
         initial_P = first(P_per),
         initial_CN = first(CN_ratio),
         initial_CP = first(CP_ratio),
         initial_NP = first(NP_ratio),
         initial_Lignin = first(Lignin_per))%>%
  #next normalizes CNP values to initial measurements for each time series and to initial masses as in Manzoni
  mutate(C_prop_initial = C_per/first(C_per),
         N_prop_initial = N_per/first(N_per),
         P_prop_initial = P_per/first(P_per),
         C_mass = ((C_per/100)*Mass_per_remaining),
         N_mass = ((N_per/100)*Mass_per_remaining),
         P_mass = ((P_per/100)*Mass_per_remaining),
         C_mass_norm = ((C_per/100)*Mass_per_remaining)/(first(C_per/100)*first(Mass_per_remaining)),
         N_mass_norm = ((N_per/100)*(Mass_per_remaining/100))/((first(N_per)/100)*first(Mass_per_remaining/100)),
         P_mass_norm = ((P_per/100)*Mass_per_remaining)/(first(P_per/100)*first(Mass_per_remaining)))%>%


  #next approximates percent mass remaining (if not provided) at each measurement day from k value
  mutate(Mass_per_remaining = case_when(is.na(Mass_per_remaining) & Remain_Mass_Category == "afdm" ~ exp(-k*Meas_Day)*100,
                                        is.na(Mass_per_remaining) & Remain_Mass_Category == "dm" ~ exp(-k*Meas_Day)*100,
                                        TRUE ~ Mass_per_remaining))%>%
  #next develops an approximation of N and P availability - imperfect, but aggregates the N and P data together about as well as possible without assuming anything else
  mutate(Water_DIN_ug.L = case_when(is.na(Water_DIN_ug.L) & is.na(Water_NH4_ug.L) ~ Water_NO3_ug.L,
                                    is.na(Water_DIN_ug.L) ~ Water_NO3_ug.L + Water_NH4_ug.L,
                                    TRUE ~ Water_DIN_ug.L),
         TN_approx = case_when(is.na(Water_TN_ug.L) ~ Water_DIN_ug.L,
                               TRUE ~ Water_TN_ug.L),
         TP_approx = case_when(is.na(Water_TP_ug.L) ~ Water_SRP_ug.L,
                               TRUE ~ Water_TP_ug.L))%>%
  #next creates a Temperature, TN, and TP average when there are multiple values given
  mutate(Temperature_C_avg = case_when(n_distinct(Temperature_C) >1 ~ mean(Temperature_C), 
                                       TRUE ~ Temperature_C),
         TN_approx_avg = case_when(n_distinct(TN_approx) >1 ~ mean(TN_approx), 
                                       TRUE ~ TN_approx),
         TP_approx_avg = case_when(n_distinct(TP_approx) >1 ~ mean(TP_approx), 
                                       TRUE ~ TP_approx))%>%
  mutate(Deg_days = Temperature_C * (Meas_Day-first(Meas_Day)))%>% #degree days
  #remove text from variable if not coarse fine or open

  mutate(mesh_cat = case_when(Mesh_Size_Category == "coarse" | Mesh_Size_Category == "fine" | Mesh_Size_Category == "open" ~ Mesh_Size_Category,
                              Mesh_Size_Category < 1 ~ "fine",
                              Mesh_Size_Category >= 1 ~ "coarse"))%>%
 
  mutate(Lotic_Lentic = case_when(System == "stream"|System == "Stream"|System == "Channel"|System == "river" ~ "Lotic",
                                  System == "wetland"|System == "Wetland"|System == "wetlands"|System == "marsh"|System == "lake"|System == "Lake"|System == "pond"|System == "reservoir" ~ "Lentic",
                                  TRUE ~ System),
         Lotic_Lentic = as_factor(Lotic_Lentic))%>%

  mutate(cum_TN = cumsum(TN_approx_avg)*Meas_Day,
         cum_TP = cumsum(TP_approx_avg)*Meas_Day)%>%
  mutate(series_index = as_factor(series_index),
         Mass_per_loss = 100-(Mass_per_remaining),
         mesh_cat = as_factor(mesh_cat),
         Remain_Mass_Category = as_factor(Remain_Mass_Category))


```

```{r trim data}
det_trim<-det%>%
  filter(Mass_per_remaining<=100)%>%
  filter(Mass_per_remaining >=20)%>%
  #filter(N_mass_norm != 0)%>%
  filter(Detritus_Type == "leaves")%>%
  filter(Detritus_Condition == "senesced")%>%
  filter(Remain_Mass_Category == "dm"|Remain_Mass_Category == "afdm")%>%
  filter(mesh_cat != "open")%>%
  mutate(Remain_Mass_Category = fct_drop(Remain_Mass_Category, only = c("dm_postleach", "afdm_postleach", "organic_carbon")))%>%
  mutate(mesh_cat = fct_drop(mesh_cat, "open"))%>%
  ungroup()

levels(det_trim$Remain_Mass_Category)
levels(det_trim$mesh_cat)
levels(det_trim$Lotic_Lentic)

```


```{r create separate lotic-lentic fine-coarse datasets}

det_loc <- det_trim%>%filter(mesh_cat == "coarse", Lotic_Lentic == "Lotic", Remain_Mass_Category == "afdm")
det_lec <- det_trim%>%filter(mesh_cat == "coarse", Lotic_Lentic == "Lentic", Remain_Mass_Category == "afdm")
det_lof <- det_trim%>%filter(mesh_cat == "fine", Lotic_Lentic == "Lotic", Remain_Mass_Category == "afdm")
det_lef <- det_trim%>%filter(mesh_cat == "fine", Lotic_Lentic == "Lentic", Remain_Mass_Category == "afdm")

```



#N mass models
##Base initial N models


```{r Ni Distro comparison}



Ni_1 <- gam(data = det_loc, N_mass_norm ~  
                     te(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = gaussian(link = "identity"),select = TRUE) 

Ni_2 <- gam(data = det_loc, N_mass_norm ~  
                     te(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = gaussian(link = "inverse"),select = TRUE) 


Ni_3 <- gam(data = det_loc, N_mass_norm ~  
                     te(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = gaussian(link = "log"),select = TRUE) 

Ni_4 <- gam(data = det_loc, N_mass_norm ~  
                     te(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "identity"),select = TRUE) 

Ni_5 <- gam(data = det_loc, N_mass_norm ~  
                     te(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "inverse"),select = TRUE) 

Ni_6 <- gam(data = det_loc, N_mass_norm ~  
                     te(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "log"),select = TRUE) 

map(list(Ni_1, Ni_2, Ni_3, Ni_4, Ni_5, Ni_6), appraise)
AIC(Ni_1, Ni_2, Ni_3, Ni_4, Ni_5, Ni_6)

# Comparisons whether interaction term is needed

Ni_5a <- gam(data = det_loc, N_mass_norm ~  
               ti(Mass_per_loss, m = 2) + ti(initial_N, m = 2)+
                     ti(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "inverse"),select = TRUE) 

Ni_5b <- gam(data = det_loc, N_mass_norm ~  
                     ti(Mass_per_loss, m = 2) + ti(initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "inverse"),select = TRUE) 

map(list(Ni_5a, Ni_5b), summary)
map(list(Ni_5a, Ni_5b), AIC)



base_loc_mod <- Ni_5 #set best model to easier name to remember for later comparisons


#plot predictions
base_loc.pred<-predict_gam(type = "link", base_loc_mod, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL, initial_N = c(0.3,0.6,1.0,1.5,2.0,3,3.5)))%>%
  mutate(initial_N = as_factor(initial_N))

#


 base_loc.p<- ggplot()+
  geom_point(data = det_loc%>%filter(Remain_Mass_Category == "afdm"), aes(Mass_per_loss, N_mass_norm), alpha = 0.3)+
    geom_smooth(data = base_loc.pred, se = FALSE, aes(x = Mass_per_loss, y = 1/fit, color = initial_N))+
    geom_abline(slope = -0.01, intercept = 1)+
    NULL
 base_loc.p
 


```
Differences in derivative change points (e.g., sooner or later) would provide evidence for immob/mineralization shifts based on initial litter decomposition. If there are no discernible differences in N peaks, that might suggest some very conserved dynamics, highly dependent on... 

```{r N Mass derivs}

N_mass_grid <- expand_grid(Mass_per_loss = seq(0,80, by = 0.2), initial_N = c(0.3,0.6,1.0,1.5,2.0,3,3.5), series_index = 578)
n_draws <- 10000
N_mass_grid_exp <- map_dfr(seq_len(n_draws), ~N_mass_grid)


gam_fitted <- fitted_samples(base_loc_mod, exclude = "s(series_index,Mass_per_loss", n = n_draws, newdata = N_mass_grid, seed = 1)%>%
  bind_cols(N_mass_grid_exp)%>%mutate(initial_N = as_factor(initial_N), draw = as_factor(draw))


deriv_samples <- gam_fitted%>%arrange(initial_N)%>%group_by(draw, initial_N)%>%mutate(fit_dif = fitted-lag(fitted))%>%mutate(Mass_per_loss = as_factor(Mass_per_loss))%>%ungroup()%>%
  group_by(initial_N, Mass_per_loss)%>%summarize(md_posterior = median(fit_dif), upper_posterior = quantile(fit_dif,0.975,na.rm = TRUE), lower_posterior = quantile(fit_dif,0.025,na.rm = TRUE))

ggplot(data = deriv_samples%>%mutate(Mass_per_loss = as.numeric(as.character(Mass_per_loss)))%>%filter(Mass_per_loss>0))+
         geom_line(aes(x= Mass_per_loss, y = md_posterior))+
  geom_line(aes(x= Mass_per_loss, y = lower_posterior), linetype = 2)+
  geom_line(aes(x= Mass_per_loss, y = upper_posterior), linetype = 2)+
         facet_wrap(.~initial_N)+
  theme_bw()+
  NULL

detderivs<-derivatives(base_loc_mod, newdata = gam.grid, term = "te(Mass_per_loss,initial_N)")
```

```{r N Mass derivs}


#create 1 pred model, then crfeate second pred model with everything offset by EPS
gam.grid<-predict_gam(type = "link", base_loc_mod, length_out = 1000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL, Remain_Mass_Category = NULL, initial_N = c(0.3,0.6,1.0,1.5,2.0,3,3.5)))%>%
  mutate(initial_N = as_factor(initial_N))%>%mutate(fit = 1/fit, se.fit = 1/se.fit)%>%
  select(-fit,-se.fit)



eps <- 1e-7
X0 <- predict(base_loc_mod, gam.grid, exclude = "s(series_index,Mass_per_loss)", type = "lpmatrix")
X1 <- gam.preds+eps
Xp <- (gam.preds_eps - gam.preds)/eps
S<-colnames(X1)
nt <- length(colnames(X1))

lD <- vector(mode = "list", length = length(colnames(X1)))
    names(lD) <- S

blah <- for(i in seq_len(nt)) {
  Xi <- Xp * 0
  want <- grep(S[i], colnames(X1))
  Xi[, want] <- Xp[, want]
  df <- Xi %*% coef(base_loc_mod)    # derivatives
  df.sd <- rowSums(Xi %*% base_loc_mod$Vp * Xi)^.5
  lD[[i]] <- list(deriv = df, se.deriv = df.sd)
}



gam.pred<-base_loc.pred
m2.d <- derivatives(base_loc_mod, term ="te(Mass_per_loss, initial_N)", newdata = gam.pred)
m2.dci <- confint(m2.d, term = Term)
m2.dsig <- signifD(gam.pred$fit, d = m2.d$derivative,
                    m2.d$upper, m2.d$lower)


plot(fit ~ Mass_per_loss, data = gam.pred, type = "n")
 #lines(fit ~ Mass_per_loss, data = gam.pred)
 #lines(fit+se.fit*1.96 ~ Mass_per_loss, data = gam.pred, lty = "dashed")
 #lines(fit-se.fit*1.96 ~ Mass_per_loss, data = gam.pred, lty = "dashed")
 lines(unlist(m2.dsig$incr) ~ Mass_per_loss, data = gam.pred, col = "blue", lwd = 3)
 lines(unlist(m2.dsig$decr) ~ Mass_per_loss, data = gam.pred, col = "red", lwd = 3)


```

#PMass analysis
```{r Pi Distro comparison}

Pi_1 <- gam(data = det_loc, P_mass_norm ~  
                        te(Mass_per_loss, initial_P, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category, m=2, bs = c("tp", "re"), full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = gaussian(link = "identity"),select = TRUE) 

Pi_2 <- gam(data = det_loc, P_mass_norm ~  
                     te(Mass_per_loss, initial_P, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category,m=2, bs = c("tp", "re"), full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = gaussian(link = "inverse"),select = TRUE) 


Pi_3 <- gam(data = det_loc, P_mass_norm ~  
                     te(Mass_per_loss, initial_P, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category,m=2, bs = c("tp", "re"), full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = gaussian(link = "log"),select = TRUE) 

Pi_4 <- gam(data = det_loc, P_mass_norm ~  
                     te(Mass_per_loss, initial_P, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category, m=2,bs = c("tp", "re"), full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "identity"),select = TRUE)
#Problems fitting Pi_4 - convergence problems

Pi_5 <- gam(data = det_loc, P_mass_norm ~  
                     te(Mass_per_loss, initial_P, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category,m=2, bs = c("tp", "re"), full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re"),
                   method = "REML", family = Gamma(link = "inverse"),select = TRUE)

det_locb <- det_loc %>% filter(initial_P < 0.25, !is.na(P_mass_norm))%>% dplyr::select(P_mass_norm, Mass_per_loss, Remain_Mass_Category, series_index, initial_P)



Pi_5b <- gam(data = det_locb, P_mass_norm ~  
                     te(Mass_per_loss, initial_P, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category,m=2, bs = c("tp", "re"), full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re"),
                   method = "REML", family = Gamma(link = "inverse"),select = TRUE)

Pi_5c <- gam(data = det_locb, list(P_mass_norm ~  
                     te(Mass_per_loss, initial_P, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category,m=2, bs = c("tp", "re"), full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re"),
                     ~te(Mass_per_loss, initial_P, m = 2)),
                   method = "REML", family = gammals(link = list( "log", "logb")),select = TRUE)


Pi_5d <- gamlss(data = det_locb, P_mass_norm ~  
                     ga(~te(Mass_per_loss, initial_P, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category,m=2, bs = c("tp", "re"), full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re"),method = "REML", select = TRUE),
                     sigma.formula = ~1, method = CG())

Pi_5e <- gamlss(data = det_locb, P_mass_norm ~  
                     ga(~te(Mass_per_loss, initial_P, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category,m=2, bs = c("tp", "re"), full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re"),method = "REML", select = TRUE),
                     sigma.formula = ~ga(~te(Mass_per_loss, initial_P, m = 2)), method = CG())



Pi_6 <- gam(data = det_loc, P_mass_norm ~  
                     te(Mass_per_loss, initial_P, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category, m=2,bs = c("tp", "re"), full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "log"),select = TRUE) 


#model comparison and evaluation
Pi_names<-list(Pi_1,Pi_2,Pi_3,Pi_4,Pi_5,Pi_6)
map(Pi_names, ~AIC(.)) #best is 5 
map(Pi_names, ~appraise(.)) #all are heteroscedastic...
summary(Pi_5) 

base_loc_mod <- Pi_5b  #set best model to easier name to remember for later comparisons


#plot predictions
base_loc.pred<-predict_gam(type = "link", base_loc_mod, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL, Remain_Mass_Category = NULL, initial_P = c(0.01,0.02,0.05,0.1,0.15,0.23)))%>%
  mutate(initial_P = as_factor(initial_P))


 base_loc.p<- ggplot()+
  geom_point(data = det_loc, aes(Mass_per_loss, P_mass_norm), alpha = 0.3)+
    geom_smooth(data = base_loc.pred, se = FALSE, aes(x = Mass_per_loss, y = 1/fit, color = initial_P))+
    geom_abline(slope = -0.01, intercept = 1)+
    NULL
 base_loc.p

```

##C:N analysis 


```{r}
CN_base1 <- gam(data = det_loc, CN_ratio ~
                        te(Mass_per_loss, initial_CN, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category, m = 2, bs = c("tp", "re"),full = TRUE)+
                        s(series_index, Mass_per_loss, bs = "re") + s(series_index, bs = "re"), 
             method = "REML", family = gaussian(link = "log"), select = TRUE)


CN_base2 <- gam(data = det_loc, CN_ratio ~ 
                        te(Mass_per_loss, initial_CN, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category, m = 2, bs = c("tp", "re"),full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re")+ s(series_index, bs = "re"), 
             method = "REML", family = Gamma(link = "log"),select = TRUE) 




AIC(CN_base1, CN_base2) #base 1 slightly favored
map(list(CN_base1,CN_base2), appraise) #2 chosen due to less heteroscedastic resids, even if deltaAIC ~90

baseCN_loc.pred<-predict_gam(type = "link", CN_base2, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL, Remain_Mass_Category = NULL, initial_CN = c(25,40,50,70,90,110,140)))%>%
  mutate(initial_CN = as_factor(initial_CN))



 base_loc_CN.p<- ggplot()+
  geom_point(data = det_loc%>%filter(!is.na(CN_ratio)), aes(Mass_per_loss, CN_ratio), alpha = 0.3)+
    geom_smooth(data = baseCN_loc.pred, se = FALSE, aes(x = Mass_per_loss, y = exp(fit), color = initial_CN))+
    geom_abline(slope = -0.01, intercept = 1)+
    #geom_ribbon(data = base_loc.pred, aes(x = Mass_per_loss, ymin = 1/(fit - 2*se.fit), ymax = 1/(fit + 2*se.fit)), alpha = 0.3)+
    NULL
 base_loc_CN.p

```


#C:P base

```{r}
CP_base1 <- gam(data = det_loc, CP_ratio ~
                        te(Mass_per_loss, initial_CP, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category, m = 2, bs = c("tp", "re"),full = TRUE)+
                        s(series_index, Mass_per_loss, bs = "re") + s(series_index, bs = "re"), 
             method = "REML", family = gaussian(link = "log"), select = TRUE)


CP_base2 <- gam(data = det_loc, CP_ratio ~ 
                        te(Mass_per_loss, initial_CP, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category, m = 2, bs = c("tp", "re"),full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re")+ s(series_index, bs = "re"), 
             method = "REML", family = Gamma(link = "log"),select = TRUE) 

#CP_base3 <- gam(data = det_loc, 
                list(CP_ratio ~ 
                        te(Mass_per_loss, initial_CP, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category, m = 2, bs = c("tp", "re"),full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re")+ s(series_index, bs = "re"),
                        ~ te(Mass_per_loss, initial_CP, m = 2)), 
             method = "REML", family = gaulss(),select = TRUE) 

CP_base4 <- gamm(data = det_loc, CP_ratio ~ 
                        te(Mass_per_loss, initial_CP, m = 2)+
                        t2(Mass_per_loss,Remain_Mass_Category, m = 2, bs = c("tp", "re"),full = TRUE) +
                        s(series_index, Mass_per_loss, bs = "re")+ s(series_index, bs = "re"),
                        weights = varIdent(~Mass_per_loss|Remain_Mass_Category), 
             method = "REML", family = gaussian(link = "log"),select = TRUE) 




AIC(CP_base1, CP_base2) #base 1 slightly favored
map(list(CP_base1,CP_base2), appraise) #2 chosen due to less heteroscedastic resids, even if deltaAIC ~90

baseCP_loc.pred<-predict_gam(type = "link", CP_base2, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL, Remain_Mass_Category = NULL, initial_CP = c(120,300,500)))%>%
  mutate(initial_CP = as_factor(initial_CP))



 base_loc_CP.p<- ggplot()+
  geom_point(data = det_loc%>%filter(!is.na(CN_ratio)), aes(Mass_per_loss, CP_ratio), alpha = 0.3)+
    geom_smooth(data = baseCP_loc.pred, se = FALSE, aes(x = Mass_per_loss, y = exp(fit), color = initial_CP))+
    geom_abline(slope = -0.01, intercept = 1)+
    #geom_ribbon(data = base_loc.pred, aes(x = Mass_per_loss, ymin = 1/(fit - 2*se.fit), ymax = 1/(fit + 2*se.fit)), alpha = 0.3)+
    NULL
 base_loc_CP.p

```