---
title: "Analyze Base NP Models"
author: "CJR"
date: "5/18/2022"
output: html_document
---
This markdown shows the analysis of the base N, P, and stoichiometry curves (how detrital nutrients evolve during decomposition based on initial nutrient content/stoichiometry)

#Setup
##Pull in tidied data and plot themes with .Rdata file; load libraries
```{r dataset for analysis}
load("C:/Users/caleb/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Code/DetNut_Data_for_Analysis.Rdata")

#Read-in libraries
library(tidyverse)
library(janitor)
library(gratia)
library(mgcv)
library(tidymv)
library(RColorBrewer)
library(cowplot)
library(ggsci)

```


#N mass models
##Base initial N models


```{r Ni curves}
#skip the line - load in selected model object and predictions
Ni_2<-readRDS("Model_Objects/Ni_base.rds")
Nbase_loc.pred<-readRDS("Model_Objects/Ni_base_preds.rds")




Ni_1 <- gam(data = det_loc, N_mass_norm ~  
                     te(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "identity"),select = TRUE) 

Ni_2 <- gam(data = det_loc, N_mass_norm ~  
                     te(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "inverse"), select = TRUE) 


Ni_3 <- gam(data = det_loc, N_mass_norm ~  
                     te(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "log"),select = TRUE) 

map(list(Ni_1, Ni_2, Ni_3), appraise) #2 is only one homoscedastic
AIC(Ni_1, Ni_2, Ni_3) #best is 2

# Comparisons whether interaction term is needed
#Setting 'select = TRUE' basically tells the GAM to perform model selection, and makes the p-values reflect whether terms are really    
# needed. Consequently, comparing plotted predictions with those from te() models above, yields different predictions thanks this 'model
# selection' - it allows the marginal effects to be penalized out similar to LASSO

Ni_2_ml <- gam(data = det_loc, N_mass_norm ~  
                     te(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "ML", family = Gamma(link = "inverse"),select = TRUE) 

Ni_2a <- gam(data = det_loc, N_mass_norm ~  
               ti(Mass_per_loss, m = 2) + ti(initial_N, m = 2)+
                     ti(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "ML", family = Gamma(link = "inverse"), select = TRUE) 

Ni_2b <- gam(data = det_loc, N_mass_norm ~  
                     ti(Mass_per_loss, m = 2) + ti(initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "ML", family = Gamma(link = "inverse"),select = TRUE) 

Ni_2c <- gam(data = det_loc, N_mass_norm ~  
                     ti(Mass_per_loss, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "ML", family = Gamma(link = "inverse"),select = TRUE) 

map(list(Ni_2a), summary) #
map(list(Ni_2a, Ni_2b), AIC) #slightly favors interactive model by 2



#plot predictions
Nbase_loc.pred<-predict_gam(type = "link", Ni_2, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL, initial_N = c(3,2,1.5,1.0,0.6,0.3)))%>%
  mutate(initial_N = as_factor(initial_N))

Nbase_loc.p<- ggplot()+
  geom_point(data = det_loc%>%filter(!is.na(N_mass_norm)&!is.na(initial_N)), aes(Mass_per_loss, N_mass_norm),shape = 1,alpha = 0.8, color = "orange")+
    geom_smooth(data = Nbase_loc.pred, se = FALSE, aes(x = Mass_per_loss, y = 1/fit, color = initial_N, linetype = initial_N), size = 1.3)+
    geom_abline(slope = -0.01, intercept = 1)+
  ylab("Normalized N Mass")+
  xlab("Mass Loss (%)")+
     scale_linetype_manual(values = c(1,10,3,4,35,6))+
  scale_color_jco()+
  guides(linetype=guide_legend(title="Initial N (%)"),color=guide_legend(title="Initial N (%)"))+
   theme_bw()+plot_theme+theme(legend.key.width = unit(1,"cm"))+
    NULL
 Nbase_loc.p
 
 
#save objects
#model object
 ###saveRDS(Ni_2,"Model_Objects/Ni_base.rds")
#predictions
###saveRDS(Nbase_loc.pred,"Model_Objects/Ni_base_preds.rds")

```


```{r N Mass derivs}

N_mass_grid <- expand_grid(Mass_per_loss = seq(0,80, by = 0.2), initial_N = c(0.3,0.6,1.0,1.5,2.0,3), series_index = 578)
n_draws <- 10000
N_mass_grid_exp <- map_dfr(seq_len(n_draws), ~N_mass_grid)


N_gam_fitted <- fitted_samples(Ni_2, exclude = "s(series_index,Mass_per_loss", n = n_draws, newdata = N_mass_grid, seed = 1)%>%
  bind_cols(N_mass_grid_exp)%>%mutate(initial_N = as_factor(initial_N), draw = as_factor(draw))


N_deriv_samples <- N_gam_fitted%>%arrange(initial_N)%>%group_by(draw, initial_N)%>%mutate(fit_dif = (fitted-lag(fitted))/0.2)%>%mutate(Mass_per_loss = as_factor(Mass_per_loss))%>%ungroup()%>%
  group_by(initial_N, Mass_per_loss)%>%summarize(md_posterior = median(fit_dif), upper_posterior = quantile(fit_dif,0.975,na.rm = TRUE), lower_posterior = quantile(fit_dif,0.025,na.rm = TRUE))
 ###saveRDS(N_deriv_samples,"Model_Objects/Ni_base_deriv.rds")

N_deriv.p<-ggplot(data = N_deriv_samples%>%mutate(Mass_per_loss = as.numeric(as.character(Mass_per_loss)))%>%filter(Mass_per_loss>0)|>
                    mutate(initial_N = fct_recode(initial_N, "0.3 %N" = "0.3","0.6 %N" ="0.6","1 %N" ="1","1.5 %N" ="1.5","2 %N" ="2","3 %N" ="3")))+
         geom_line(aes(x= Mass_per_loss, y = md_posterior), size = 1)+
  geom_line(aes(x= Mass_per_loss, y = lower_posterior), linetype = 2)+
  geom_line(aes(x= Mass_per_loss, y = upper_posterior), linetype = 2)+
         facet_wrap(.~initial_N)+
  ylab(expression(atop(bold("Rate of Change"), bold("(N/Mass loss)"))))+
  xlab("Mass Loss (%)")+
  geom_hline(yintercept = 0)+
  theme_bw()+deriv_plot_theme+
  NULL
N_deriv.p
```

```{r cowplot N mass and deriv plots}

N_grid <- plot_grid(Nbase_loc.p,N_deriv.p, nrow = 2, labels = "AUTO", label_size = 20)
ggsave2("C:/Users/caleb/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Figures/DetNut_Fig2_N_mass.pdf", N_grid, width = 7.8, height = 9.1, dpi = 600)

```

#PMass analysis
```{r Pi curves}
#Skip the line - load in model object and predictions:
#Pi_2<-readRDS("Model_Objects/Pi_base.rds")

#Pbase_loc.pred<-readRDS("Model_Objects/Pi_base_preds.rds")


Pi_1 <- gam(data = det_loc, P_mass_norm ~  
                     te(Mass_per_loss, initial_P, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "identity"),select = TRUE)

Pi_2 <- gam(data = det_loc, P_mass_norm ~  
                     te(Mass_per_loss, initial_P, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"),
                   method = "REML", family = Gamma(link = "inverse"),select = TRUE)


Pi_3 <- gam(data = det_loc, P_mass_norm ~  
                     te(Mass_per_loss, initial_P, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "REML", family = Gamma(link = "log")) 



#model comparison and evaluation
Pi_names<-list(Pi_1,Pi_2,Pi_3) 
map(Pi_names, ~AIC(.)) #best is 2
map(Pi_names, ~appraise(.)) #all are heteroscedastic, with no great options, 2 OK
summary(Pi_2) 

Pi_2a <- gam(data = det_loc, P_mass_norm ~  
               ti(Mass_per_loss, m = 2) + ti(initial_P, m = 2)+
                     ti(Mass_per_loss, initial_P, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "ML", family = Gamma(link = "inverse"),select = TRUE) 

Pi_2b <- gam(data = det_loc, P_mass_norm ~  
                     ti(Mass_per_loss, m = 2) + ti(initial_P, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "ML", family = Gamma(link = "inverse"),select = TRUE) 

Pi_2c <- gam(data = det_loc, P_mass_norm ~  
                     ti(Mass_per_loss, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re"), 
                   method = "ML", family = Gamma(link = "inverse"),select = TRUE) 

map(list(Pi_2a, Pi_2b), summary) #interaction term pval = 0.0012
map(list(Pi_2a, Pi_2b,Pi_2c), AIC) #AIC -11.2, favoring mass loss model



#plot predictions
Pbase_loc.pred<-predict_gam(type = "link", Pi_2, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL, Remain_Mass_Category = NULL, initial_P = c(0.01,0.02,0.05,0.1,0.15,0.23)))%>%
  mutate(initial_P = as_factor(initial_P))


 Pbase_loc.p<- ggplot()+
  geom_point(data = det_loc%>%filter(!is.na(P_mass_norm)&!(is.na(initial_P))), aes(Mass_per_loss, P_mass_norm), shape = 1,alpha = 0.8, color = "orange")+
    geom_smooth(data = Pbase_loc.pred, se = FALSE, aes(x = Mass_per_loss, y = 1/fit, color = initial_P, linetype = initial_P), size = 1.5)+
    geom_abline(slope = -0.01, intercept = 1)+
   xlab("Mass Loss (%)")+
   ylab("Normalized P mass")+
    scale_linetype_manual(values = c(1,10,3,4,35,6))+
  scale_color_jco()+
  guides(linetype=guide_legend(title="Initial P (%)"),color=guide_legend(title="Initial P (%)"))+
   theme_bw()+plot_theme+theme(legend.key.width = unit(1,"cm"))+
    NULL
 Pbase_loc.p

 
#save mod objects and predictions
 ##saveRDS(Pi_2,"Model_Objects/Pi_base.rds")

###saveRDS(Pbase_loc.pred,"Model_Objects/Pi_base_preds.rds")
```

```{r P mass derivs}
P_mass_grid <- expand_grid(Mass_per_loss = seq(0,80, by = 0.2), initial_P = c(0.01,0.02,0.05,0.1,0.15,0.23), series_index = 578)
n_draws <- 10000
P_mass_grid_exp <- map_dfr(seq_len(n_draws), ~P_mass_grid)


P_gam_fitted <- fitted_samples(Pi_2, exclude = "s(series_index,Mass_per_loss", n = n_draws, newdata = P_mass_grid, seed = 1)%>%
  bind_cols(P_mass_grid_exp)%>%mutate(initial_P = as_factor(initial_P), draw = as_factor(draw))


P_deriv_samples <- P_gam_fitted%>%arrange(initial_P)%>%group_by(draw, initial_P)%>%mutate(fit_dif = (fitted-lag(fitted))/0.2)%>%mutate(Mass_per_loss = as_factor(Mass_per_loss))%>%ungroup()%>%
  group_by(initial_P, Mass_per_loss)%>%summarize(md_posterior = median(fit_dif), upper_posterior = quantile(fit_dif,0.975,na.rm = TRUE), lower_posterior = quantile(fit_dif,0.025,na.rm = TRUE))

###saveRDS(P_deriv_samples,"Model_Objects/Pi_base_deriv.rds")

P_deriv.p<-ggplot(data = P_deriv_samples%>%mutate(Mass_per_loss = as.numeric(as.character(Mass_per_loss)))%>%filter(Mass_per_loss>0)|>
                    mutate(initial_P = fct_recode(initial_P, "0.01 %P" = "0.01","0.02 %P" ="0.02","0.05 %P" ="0.05","0.1 %P" ="0.1","0.15 %P" ="0.15","0.23 %P" ="0.23")))+
         geom_line(aes(x= Mass_per_loss, y = md_posterior), size = 1)+
  geom_line(aes(x= Mass_per_loss, y = lower_posterior), linetype = 2)+
  geom_line(aes(x= Mass_per_loss, y = upper_posterior), linetype = 2)+
         facet_wrap(.~initial_P)+
  geom_hline(yintercept = 0)+
  ylab(expression(atop(bold("Rate of Change"), bold("(P/Mass loss)"))))+
  xlab("Mass Loss (%)")+
  theme_bw()+deriv_plot_theme+
  NULL
P_deriv.p
```

```{r cowplot P mass and deriv plots}

P_grid<-plot_grid(Pbase_loc.p,P_deriv.p, nrow = 2, labels = "AUTO", label_size = 20)
ggsave2("C:/Users/caleb/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Figures/DetNut_Fig3_P_mass.pdf", P_grid, width = 7.8, height = 9.1, dpi = 600)
```

```{r cowplot N and P plots}
N_and_P_grid<-plot_grid(Nbase_loc.p,N_deriv.p,Pbase_loc.p,P_deriv.p, nrow = 2, labels = "AUTO", label_size = 20)


ggsave2("C:/Users/caleb/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Figures/DetNut_Fig2_NP_mass.pdf", N_and_P_grid, width = 14, height = 9.1, dpi = 600)

```

##C:N analysis 


```{r CN curves}

CN_base1 <- gam(data = det_loc_cnp, CN_ratio ~ 
                        te(Mass_per_loss, initial_CN, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+ s(series_index, bs = "re"), 
             method = "REML", family = Gamma(link = "log"), select = TRUE) 


map(list(CN_base1), appraise) #acceptable heteroscedasticity

CN_1a <- gam(data = det_loc_cnp, CN_ratio ~  
               ti(Mass_per_loss, m = 2) + ti(initial_CN, m = 2)+
                     ti(Mass_per_loss, initial_CN, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "log"),select = TRUE) 

CN_1b <- gam(data = det_loc_cnp, CN_ratio ~  
                     ti(Mass_per_loss, m = 2) + ti(initial_CN, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "log"),select = TRUE) 

map(list(CN_1a, CN_1b), summary) #interaction term pval = 2e-16
map(list(CN_1a, CN_1b), AIC) #AIC favors interactive model

baseCN_loc.pred<-predict_gam(type = "link", CN_base1, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL, Remain_Mass_Category = NULL, initial_CN = c(140,100,70,50,40,25)))%>%
  mutate(initial_CN = fct_relevel(as_factor(initial_CN), c("140","100","70","50","40","25")))


 CNbase_loc.p<- ggplot()+
  geom_point(data = det_loc_cnp%>%filter(!is.na(CN_ratio)&!is.na(initial_CN)), aes(Mass_per_loss, CN_ratio), color = "orange",alpha = 0.8, shape = 1)+
    geom_smooth(data = baseCN_loc.pred, se = FALSE, aes(x = Mass_per_loss, y = exp(fit), color = initial_CN, linetype = initial_CN), size = 1.5)+
    geom_abline(slope = -0.01, intercept = 1)+
   scale_y_continuous(breaks = c(20,50,100,150))+
    ylab("C:N")+
   xlab("Mass Loss (%)")+
    scale_linetype_manual(values = c(1,10,3,4,35,6))+
  scale_color_jco()+
  guides(linetype=guide_legend(title="Initial C:N"),color=guide_legend(title="Initial C:N"))+
   theme_bw()+plot_theme+theme(legend.key.width = unit(1,"cm"))+
    NULL
   
   CNbase_loc.p

   
 #save mod objects and predictions
 ##saveRDS(CN_base1,"Model_Objects/CN_base.rds")

###saveRDS(baseCN_loc.pred,"Model_Objects/CN_base_preds.rds") 
```

```{r C:N derivs}
CN_mass_grid <- expand_grid(Mass_per_loss = seq(0,80, by = 0.2), initial_CN = c(25,40,50,70,100,140), series_index = 578)
n_draws <- 10000
CN_mass_grid_exp <- map_dfr(seq_len(n_draws), ~CN_mass_grid)


CN_gam_fitted <- fitted_samples(CN_base1, exclude = "s(series_index,Mass_per_loss", n = n_draws, newdata = CN_mass_grid, seed = 1)%>%
  bind_cols(CN_mass_grid_exp)%>%mutate(initial_CN = as_factor(initial_CN), draw = as_factor(draw), fitted = log(fitted))
#log transformed fitted values because *ratios*


CN_deriv_samples <- CN_gam_fitted%>%arrange(initial_CN)%>%group_by(draw, initial_CN)%>%mutate(fit_dif = (fitted-lag(fitted))/0.2)%>%mutate(Mass_per_loss = as_factor(Mass_per_loss))%>%ungroup()%>%
  group_by(initial_CN, Mass_per_loss)%>%summarize(md_posterior = median(fit_dif), upper_posterior = quantile(fit_dif,0.975,na.rm = TRUE), lower_posterior = quantile(fit_dif,0.025,na.rm = TRUE))

###saveRDS(CN_deriv_samples,"Model_Objects/CN_base_deriv.rds")

CN_deriv.p<-ggplot(data = CN_deriv_samples%>%mutate(Mass_per_loss = as.numeric(as.character(Mass_per_loss)))%>%filter(Mass_per_loss>0))+
         geom_line(aes(x= Mass_per_loss, y = exp(md_posterior)), size = 1)+
  geom_line(aes(x= Mass_per_loss, y = exp(lower_posterior)), linetype = 2)+
  geom_line(aes(x= Mass_per_loss, y = exp(upper_posterior)), linetype = 2)+
  scale_y_continuous(trans="log")+
         facet_wrap(.~initial_CN)+
  geom_hline(yintercept = 1)+
  ylab(expression(atop(bold("Rate of Change"), bold("(C:N/Mass loss)"))))+
  xlab("Mass Loss (%)")+
  theme_bw()+deriv_plot_theme+
  NULL
CN_deriv.p
```

```{r cowplot C:N and deriv plots}

CN_grid<-plot_grid(CNbase_loc.p,CN_deriv.p, nrow = 2, labels = "AUTO", label_size = 20)
ggsave2("C:/Users/caleb/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Figures/DetNut_Fig4_CN.pdf", CN_grid, width = 7.8, height = 9.1, dpi = 600)
```

#C:P base
CP_base1<-readRDS("Model_Objects/CP_base.rds")

```{r CP Mass loss}

CP_base1 <- gam(data = det_loc_cnp, CP_ratio ~
                  te(Mass_per_loss, initial_CP, m= 2)+
                 s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"),
                method = "REML", family = Gamma(link = "log"), select = TRUE)


map(list(CP_base1), appraise) #kind of weird, but spread seems not too big of a problem

CP_1a <- gam(data = det_loc_cnp, CP_ratio ~  
               ti(Mass_per_loss, m = 2) + ti(initial_CP, m = 2)+
                     ti(Mass_per_loss, initial_CP, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "log"),select = TRUE) 

CP_1b <- gam(data = det_loc_cnp, CP_ratio ~  
                     ti(Mass_per_loss, m = 2) + ti(initial_CP, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "log"),select = TRUE) 

map(list(CP_1a, CP_1b), summary) #interaction term highly significant - 0.0003
map(list(CP_1a, CP_1b), AIC) #favors interaction model - delAIC 45.1


baseCP_loc.pred<-predict_gam(type = "link", CP_base1, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL, Remain_Mass_Category = NULL, initial_CP = c(5000,3000,2000,1600,1200,800)))%>%
  mutate(initial_CP = fct_relevel(as_factor(initial_CP), c("5000","3000","2000","1600","1200","800")))



 CPbase_loc.p<- ggplot()+
  geom_point(data = det_loc_cnp%>%filter(!is.na(CP_ratio)&!is.na(initial_CP)), aes(Mass_per_loss, CP_ratio), alpha = 0.8, color = "orange", shape = 1)+
    geom_smooth(data = baseCP_loc.pred, se = FALSE, aes(x = Mass_per_loss, y = exp(fit), color = initial_CP, linetype = initial_CP), size = 1.5)+
   scale_y_continuous(breaks = c(1000,3000,5000, 8000))+
      ylab("C:P")+
   xlab("Mass loss (%)")+
       scale_linetype_manual(values = c(1,10,3,4,35,6))+
  scale_color_jco()+
  guides(linetype=guide_legend(title="Initial C:P"),color=guide_legend(title="Initial C:P"))+
   theme_bw()+plot_theme+theme(legend.key.width = unit(1,"cm"))+
    NULL
 CPbase_loc.p

 
   
 #save mod objects and predictions
##saveRDS(CP_base1,"Model_Objects/CP_base.rds")
 ##saveRDS(baseCP_loc.pred,"Model_Objects/CP_base_preds.rds")
```

CP_deriv_samples <- readRDS("Model_Objects/CP_base_deriv.rds")
```{r CP derivs}
CP_mass_grid <- expand_grid(Mass_per_loss = seq(0,80, by = 0.2), initial_CP = c(800,1200,1600,2000,3000,5000), series_index = 578)
n_draws <- 10000
CP_mass_grid_exp <- map_dfr(seq_len(n_draws), ~CP_mass_grid)


CP_gam_fitted <- fitted_samples(CP_base1, exclude = "s(series_index,Mass_per_loss", n = n_draws, newdata = CP_mass_grid, seed = 1)%>%
  bind_cols(CP_mass_grid_exp)%>%mutate(initial_CP = as_factor(initial_CP), draw = as_factor(draw), fitted = log(fitted))
#log transformed fitted values because *ratios*


CP_deriv_samples <- CP_gam_fitted%>%arrange(initial_CP)%>%group_by(draw, initial_CP)%>%mutate(fit_dif = (fitted-lag(fitted))/0.2)%>%mutate(Mass_per_loss = as_factor(Mass_per_loss))%>%ungroup()%>%
  group_by(initial_CP, Mass_per_loss)%>%summarize(md_posterior = median(fit_dif), upper_posterior = quantile(fit_dif,0.975,na.rm = TRUE), lower_posterior = quantile(fit_dif,0.025,na.rm = TRUE))


###saveRDS(CP_deriv_samples,"Model_Objects/CP_base_deriv.rds")

CP_deriv.p<-ggplot(data = CP_deriv_samples%>%mutate(Mass_per_loss = as.numeric(as.character(Mass_per_loss)))%>%filter(Mass_per_loss>0))+
         geom_line(aes(x= Mass_per_loss, y = exp(md_posterior)),size = 1)+
  geom_line(aes(x= Mass_per_loss, y = exp(lower_posterior)), linetype = 2)+
  geom_line(aes(x= Mass_per_loss, y = exp(upper_posterior)), linetype = 2)+
         facet_wrap(.~initial_CP)+
  geom_hline(yintercept = 1)+
  scale_y_continuous(trans="log")+
  ylab(expression(atop(bold("Rate of Change"), bold("(C:P/Mass loss)"))))+
  xlab("Mass Loss (%)")+
  theme_bw()+deriv_plot_theme+
  NULL
CP_deriv.p
```

```{r cowplot C:P and deriv plots}

CP_grid<-plot_grid(CPbase_loc.p,CP_deriv.p, nrow = 2, labels = "AUTO", label_size = 20)
ggsave2("C:/Users/caleb/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Figures/DetNut_Fig5_CP.pdf", CP_grid, width = 7.8, height = 9.1, dpi = 600)
```

```{r cowplot C:N and C:P plots}
CN_and_CN_grid<-plot_grid(CNbase_loc.p,CN_deriv.p, CPbase_loc.p,CP_deriv.p, nrow = 2, labels = "AUTO", label_size = 20)

ggsave2("C:/Users/caleb/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Figures/DetNut_Fig3_CNP.pdf", CN_and_CN_grid, width = 14, height = 9.1, dpi = 600)

```
#N:P base

```{r NP mass loss curves}


NP_base1 <- gam(data = det_loc_cnp, NP_ratio ~
                  te(Mass_per_loss, initial_NP, m= 2)+
                  s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"),
                method = "REML", family = Gamma(link = "log"), select = TRUE)

map(list(NP_base1), appraise) #heteroscedasticity not too bad  

NP_1a <- gam(data = det_loc_cnp, NP_ratio ~  
               ti(Mass_per_loss, m = 2) + ti(initial_NP, m = 2)+
                     ti(Mass_per_loss, initial_NP, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "log"),select = TRUE) 

NP_1b <- gam(data = det_loc_cnp, NP_ratio ~  
                     ti(Mass_per_loss, m = 2) + ti(initial_NP, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "log"),select = TRUE) 

map(list(NP_1a, NP_1b), summary) #interaction term - 0.017
map(list(NP_1a, NP_1b), AIC) #delAIC 8.06 in favor of interactive model
1

baseNP_loc.pred <-predict_gam(type = "link", NP_base1, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL, Remain_Mass_Category = NULL, initial_NP = c(100,70,55,40,25,10)))%>%
  mutate(initial_NP = fct_relevel(as_factor(initial_NP), c("100","70","55","40","25","10")))



 NPbase_loc.p<- ggplot()+
  geom_point(data = det_loc_cnp%>%filter(!is.na(NP_ratio)&!is.na(initial_NP)), aes(Mass_per_loss, NP_ratio), color = "orange",alpha = 0.8, shape = 1)+
    geom_smooth(data = baseNP_loc.pred, se = FALSE, aes(x = Mass_per_loss, y = exp(fit), color = initial_NP, linetype = initial_NP), size = 1.5)+
   scale_y_continuous(breaks = c(10,50,100, 200))+
   ylab("N:P")+
   xlab("Mass loss (%)")+
       scale_linetype_manual(values = c(1,10,3,4,35,6))+
  scale_color_jco()+
  guides(linetype=guide_legend(title="Initial N:P"),color=guide_legend(title="Initial N:P"))+
   theme_bw()+plot_theme+theme(legend.key.width = unit(1,"cm"))+
    NULL
 NPbase_loc.p
 
  #save mod objects and predictions
 ##saveRDS(NP_base1,"Model_Objects/NP_base.rds")

###saveRDS(baseNP_loc.pred,"Model_Objects/NP_base_preds.rds")

```

```{r NP derivs}
NP_mass_grid <- expand_grid(Mass_per_loss = seq(0,80, by = 0.2), initial_NP = c(100,70,55,40,25,10), series_index = 578)
n_draws <- 10000
NP_mass_grid_exp <- map_dfr(seq_len(n_draws), ~NP_mass_grid)


NP_gam_fitted <- fitted_samples(NP_base1, exclude = "s(series_index,Mass_per_loss", n = n_draws, newdata = NP_mass_grid, seed = 1)%>%
  bind_cols(NP_mass_grid_exp)%>%mutate(initial_NP = as_factor(initial_NP), draw = as_factor(draw), fitted = log(fitted))
#log transformed fitted values because *ratios*


NP_deriv_samples <- NP_gam_fitted%>%arrange(initial_NP)%>%group_by(draw, initial_NP)%>%mutate(fit_dif = (fitted-lag(fitted))/0.2)%>%mutate(Mass_per_loss = as_factor(Mass_per_loss))%>%ungroup()%>%
  group_by(initial_NP, Mass_per_loss)%>%summarize(md_posterior = median(fit_dif), upper_posterior = quantile(fit_dif,0.975,na.rm = TRUE), lower_posterior = quantile(fit_dif,0.025,na.rm = TRUE))

###saveRDS(NP_deriv_samples,"Model_Objects/NP_base_deriv.rds")

NP_deriv.p <- ggplot(data = NP_deriv_samples%>%mutate(Mass_per_loss = as.numeric(as.character(Mass_per_loss)))%>%filter(Mass_per_loss>0))+
         geom_line(aes(x= Mass_per_loss, y = exp(md_posterior)),size = 1)+
  geom_line(aes(x= Mass_per_loss, y = exp(lower_posterior)), linetype = 2)+
  geom_line(aes(x= Mass_per_loss, y = exp(upper_posterior)), linetype = 2)+
         facet_wrap(.~initial_NP)+
  geom_hline(yintercept = 1)+
  scale_y_continuous(trans="log")+
  ylab(expression(atop(bold("Rate of Change"), bold("(N:P/Mass loss)"))))+
  xlab("Mass Loss (%)")+
  theme_bw()+deriv_plot_theme+
  NULL
NP_deriv.p
```

```{r cowplot C:P and deriv plots}

NP_grid<-plot_grid(NPbase_loc.p,NP_deriv.p, nrow = 2, labels = "AUTO", label_size = 20)
ggsave2("C:/Users/caleb/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Figures/DetNut_Fig6_NP.pdf", NP_grid, width = 7.8, height = 9.1, dpi = 600)
```

