---
title: "NP_Exploratory"
author: "CJR"
date: '2022-06-14'
output: html_document
---
This markdown starts with base models found in NP_Base_Models.Rmd markdown to explore how covariates might influence the base curves. 

#Setup

Load-in the tidied dataset and packages
```{r setup, include=FALSE}
#load("C:/Users/caleb/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Code/DetNut_Data_for_Analysis.Rdata")

#Read-in libraries
library(tidyverse)
library(janitor)
library(gratia)
library(mgcv)
library(tidymv)
library(RColorBrewer)

```


Need to consider covariate structure
```{r}
ggplot(data=det_loc, aes(x = Temperature_C_avg))+geom_histogram() #Temp is fairly evenly distributed
ggplot(data=det_loc, aes(x = Velocity_m.s))+geom_histogram() #okay coverage, but low quantity
ggplot(data=det_loc, aes(x = pH))+geom_histogram()

ggplot(data=det_loc, aes(x = TN_approx))+geom_histogram() #relatively few data sit above 7500 ug/L, where the rest of the dataset is mostly below 2500 ug/L - it would be appropriate to remove those as they could be strong leveraging points but represent a very small portion of the dataset.
det_loc_tn <- det_loc%>%filter(TN_approx<7500)

ggplot(data=det_loc, aes(x = TP_approx))+geom_histogram() #Again, one grouping of data is very high relative to the rest of the dataset, and could be strong leveraging points, so we will remove those
det_loc_tp <- det_loc%>%filter(TP_approx <2000)

det_loc<-det_loc%>%mutate(N_mass = N_per*Mass_per_remaining)

```


#N mass exploratory models


```{r Ni x Temp}

Ni_base <- gam(data = det_loc, N_mass ~  
               ti(Mass_per_loss, m = 2) + ti(initial_N, m = 2)+
                     ti(Mass_per_loss, initial_N, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "REML", family = Gamma(link = "inverse"),select = TRUE) 

Ni_temp_ml <- gam(data = det_loc, N_mass ~  
                  ti(Mass_per_loss, m = 2) + ti(initial_N, m = 2)+
                  ti(Mass_per_loss, initial_N, m = 2)+
               ti(Mass_per_loss,initial_N, Temperature_C_avg, m = 1)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "inverse"),select = TRUE) 

Ni_temp <- gam(data = det_loc, N_mass ~  
                  ti(Mass_per_loss, m = 2) + ti(initial_N, m = 2)+
                  ti(Mass_per_loss, initial_N, m = 2)+
               ti(Mass_per_loss,initial_N, Temperature_C_avg, m = 1)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "REML", family = Gamma(link = "inverse"),select = TRUE) 

 ###saveRDS(Ni_temp,"Model_Objects/Ni_temp.rds")

AIC(Ni_temp_ml,Ni_base)
summary(Ni_temp_ml)
appraise(Ni_temp)


Ni_T.pred<-predict_gam(type = "link", Ni_temp, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL,Remain_Mass_Category = NULL, initial_N = c(3,2.0,1.5,1.0,0.6,0.3),Temperature_C_avg = c(5,10,15,20)))%>%
   mutate(Temperature_C_avg = as_factor(Temperature_C_avg), initial_N = as_factor(initial_N))
  


 Ni_T.p<- ggplot()+
    geom_smooth(data = Ni_T.pred, se = FALSE, aes(x = Mass_per_loss, y = 1/fit, color = Temperature_C_avg, linetype = Temperature_C_avg))+
   facet_grid(initial_N~., scales = "free")+
   geom_ribbon(data = Ni_T.pred, aes(x = Mass_per_loss, ymin = 1/(fit - 2*se.fit), ymax = 1/(fit + 2*se.fit)), alpha = 0.3)+
    geom_abline(slope = -0.01, intercept = 1)+
    NULL
 Ni_T.p
 



```

```{r Ni x pH}


Ni_ph_ml <- gam(data = det_loc, N_mass ~  
                  ti(Mass_per_loss, m = 2) + ti(initial_N, m = 2)+
                  ti(Mass_per_loss, initial_N, m = 2)+
               ti(Mass_per_loss,initial_N, pH, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "inverse"),select = TRUE) 

###"Model has more coefficiencts than data"

```

```{r Ni x TN}


Ni_tn_ml <- gam(data = det_loc, N_mass ~  
                  ti(Mass_per_loss, m = 2) + ti(initial_N, m = 2)+
                  ti(Mass_per_loss, initial_N, m = 2)+
                  ti(Mass_per_loss,initial_N, TN_approx, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "inverse"),select = TRUE) 

Ni_tn <- gam(data = det_loc, N_mass ~  
                  ti(Mass_per_loss, m = 2) + ti(initial_N, m = 2)+
                  ti(Mass_per_loss, initial_N, m = 2)+
                  ti(Mass_per_loss,initial_N, TN_approx, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "REML", family = Gamma(link = "inverse"),select = TRUE) 



summary(Ni_tn_ml)
appraise(Ni_tn)
###saveRDS(Ni_tn,"Model_Objects/Ni_tn.rds")

Ni_TN.pred<-predict_gam(type = "link", Ni_tn, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL,Remain_Mass_Category = NULL, initial_N = c(3,2,1.5,1.0,0.6,0.3),TN_approx = c(300,600,1000,1500)))%>%
   mutate(TN_approx = as_factor(TN_approx), initial_N = as_factor(initial_N))
  


 Ni_TN.p<- ggplot()+
    geom_smooth(data = Ni_TN.pred, se = FALSE, aes(x = Mass_per_loss, y = 1/fit, color = TN_approx, linetype = TN_approx))+
   facet_grid(initial_N~., scales = "free")+
   geom_ribbon(data = Ni_TN.pred, aes(x = Mass_per_loss, ymin = 1/(fit - 2*se.fit), ymax = 1/(fit + 2*se.fit)), alpha = 0.3)+
    geom_abline(slope = -0.01, intercept = 1)+
    NULL
 Ni_TN.p
```



```{r Ni x TP}

Ni_tp_ml <- gam(data = det_loc, N_mass ~  
                  ti(Mass_per_loss) + ti(initial_N)+
                  ti(Mass_per_loss, initial_N, m = 2)+
                  ti(Mass_per_loss,initial_N, TP_approx, m = 1)+ 
               s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "ML", family = gaussian(link = "inverse"),select = TRUE)

Ni_tp <- gam(data = det_loc, N_mass ~  
                  ti(Mass_per_loss) + ti(initial_N)+
                  ti(Mass_per_loss, initial_N, m = 2)+
                  ti(Mass_per_loss,initial_N, TP_approx, m = 1)+ 
               s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "REML", family = gaussian(link = "inverse"),select = TRUE)



summary(Ni_tp_ml)
appraise(Ni_tp)

###saveRDS(Ni_tp,"Model_Objects/Ni_tp.rds")

Ni_TP.pred<-predict_gam(type = "link", Ni_tp, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL,Remain_Mass_Category = NULL, initial_N = c(3,2,1.5,1.0,0.6,0.3),TP_approx = c(50,100,300,500)))%>%
   mutate(TP_approx = as_factor(TP_approx), initial_N = as_factor(initial_N))
  


 Ni_TP.p<- ggplot()+
    geom_smooth(data = Ni_TP.pred, se = FALSE, aes(x = Mass_per_loss, y = 1/fit, color = initial_N, linetype = TP_approx))+
   facet_grid(initial_N~TP_approx, scales = "free")+
   geom_ribbon(data = Ni_TP.pred, aes(x = Mass_per_loss, ymin = 1/(fit - 2*se.fit), ymax = 1/(fit + 2*se.fit)), alpha = 0.3)+
    geom_abline(slope = -0.01, intercept = 1)+
    NULL
 Ni_TP.p
 
```

#CN exploratory models

```{r CN X Temp}
CN_base1 <- gam(data = det_loc_cnp, CN_ratio ~
                        te(Mass_per_loss, initial_CN, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re") + s(series_index, bs = "re"), 
             method = "REML", family = Gamma(link = "log"), select = TRUE)


CN_temp_ml <- gam(data = det_loc_cnp, CN_ratio ~  
                  ti(Mass_per_loss) + ti(initial_CN)+
                  ti(Mass_per_loss, initial_CN)+
               ti(Mass_per_loss,initial_CN, Temperature_C_avg, m = 1)+
                        s(series_index, Mass_per_loss, bs = "re")+ s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "log"),select = TRUE) 


CN_temp <- gam(data = det_loc_cnp, CN_ratio ~  
                  ti(Mass_per_loss) + ti(initial_CN)+
                  ti(Mass_per_loss, initial_CN)+
               ti(Mass_per_loss,initial_CN, Temperature_C_avg, m = 1)+
                        s(series_index, Mass_per_loss, bs = "re")+ s(series_index, bs = "re"), 
                   method = "REML", family = Gamma(link = "log"),select = TRUE) 

AIC(CN_base1,CN_temp_ml)
summary(CN_temp_ml)
appraise(CN_temp)

###saveRDS(CN_temp,"Model_Objects/CN_temp.rds")
###CN_temp <- readRDS("Model_Objects/CN_temp.rds")

CN_T.pred<-predict_gam(type = "link", CN_temp, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL,Remain_Mass_Category = NULL, initial_CN = c(140,100,70,50,40,25),Temperature_C_avg = c(7.5,10,15,20)))%>%
   mutate(Temperature_C_avg = as_factor(Temperature_C_avg), initial_CN = as_factor(initial_CN))
  

 CN_T.p<- ggplot(data = CN_T.pred%>%mutate(Temperature_C_avg = fct_recode(Temperature_C_avg, "7.5 C" = "7.5", "10 C" = "10","15 C" = "15", "20 C" = "20")))+
    geom_smooth(se = FALSE, aes(x = Mass_per_loss, y = exp(fit), color = Temperature_C_avg, linetype = Temperature_C_avg))+
   facet_grid(initial_CN~Temperature_C_avg, scales = "free")+
   geom_ribbon(aes(x = Mass_per_loss, ymin = exp(fit - 2*se.fit), ymax = exp(fit + 2*se.fit)), alpha = 0.3)+
   ylab("C:N (molar)")+
   xlab("Mass loss (%)")+ggtitle("Temperature x C:N")+theme(legend.position ="none")+
  NULL
 CN_T.p

```


```{r categorized temp attempt}
#visualize and bin temperature
det_cnp_temp<- det_loc_cnp|>
  mutate(Temp_cat = as_factor(case_when(Temperature_C_avg<=10 ~ "Low",
                              Temperature_C_avg>10 &Temperature_C_avg<=18 ~ "Med",
                              Temperature_C_avg>18 ~ "High")))|>
  filter(!is.na(Temp_cat))


  CN_temp <- gam(data = det_cnp_temp, CN_ratio ~  
                  ti(Mass_per_loss, m = 2, k = 8) + ti(initial_CN, m = 2)+
                  ti(Mass_per_loss, initial_CN, m = 2)+
               ti(Mass_per_loss,initial_CN, by=Temp_cat)+
                        s(series_index, Mass_per_loss, bs = "re")+ s(series_index, bs = "re"), 
                   method = "REML", family = Gamma(link = "log"),select = TRUE) 
 
   CN_T.pred<-predict_gam(type = "link", CN_temp, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL,Remain_Mass_Category = NULL, initial_CN = c(140,100,70,50,40,25),Temp_cat = c("Low","Med", "High")))%>%
   mutate(Temp_cat = as_factor(Temp_cat), initial_CN = as_factor(initial_CN))
  

 CN_T.p<- ggplot(data = CN_T.pred)+
    geom_smooth(se = FALSE, aes(x = Mass_per_loss, y = exp(fit), color = Temp_cat, linetype = Temp_cat))+
   facet_grid(initial_CN~Temp_cat, scales = "free")+
   geom_ribbon(aes(x = Mass_per_loss, ymin = exp(fit - 2*se.fit), ymax = exp(fit + 2*se.fit)), alpha = 0.3)+
   ylab("C:N (molar)")+
   xlab("Mass loss (%)")+ggtitle("Temperature x C:N")+theme(legend.position ="none")+
  NULL
 CN_T.p

```
Temp significant change in curves - plot derivatives. 

This takes a long time run fitted_samples() with 5000 draws (maybe 15 minutes or more), so the object important in running the plot has been saved as an .rds file.

This can be read in with CNxTemp_deriv_samples <- readRDS("Model_Objects/CN_temp_derivs.rds") where file connection is set to correct working directory where .rds file is stored.
```{r Temp covar curve derivatives}

CNxTemp_mass_grid <- expand_grid(Mass_per_loss = seq(0,80, by = 0.2), Temperature_C_avg = c(7.5,10,15,20),initial_CN = c(25,40,50,70,100,140), series_index = 4)
n_draws <- 500
CNxTemp_mass_grid_exp <- map_dfr(seq_len(n_draws), ~CNxTemp_mass_grid)


CNxTemp_gam_fitted <- fitted_samples(CN_temp, exclude = c("s(series_index)", "s(series_index, Mass_per_loss)"), n = n_draws, newdata = CNxTemp_mass_grid, seed = 1, scale = "response")%>%
  bind_cols(CNxTemp_mass_grid_exp)%>%mutate(initial_CN = as_factor(initial_CN), Temperature_C_avg = as_factor(Temperature_C_avg), draw = as_factor(draw), fitted = log(fitted))
#log transformed fitted values because *ratios*


CNxTemp_deriv_samples <- CNxTemp_gam_fitted%>%arrange(initial_CN, Temperature_C_avg)%>%group_by(draw, initial_CN,Temperature_C_avg)%>%mutate(fit_dif = (fitted-lag(fitted))/0.2)%>%mutate(Mass_per_loss = as_factor(Mass_per_loss))%>%ungroup()%>%
  group_by(initial_CN,Temperature_C_avg, Mass_per_loss)%>%summarize(md_posterior = median(fit_dif), upper_posterior = quantile(fit_dif,0.975,na.rm = TRUE), lower_posterior = quantile(fit_dif,0.025,na.rm = TRUE))

#export object because it took about 15 minutes to create with 5000 draws...
#saveRDS(CNxTemp_deriv_samples,"Model_Objects/CN_temp_derivs.rds")


CNxTemp_deriv.p<-ggplot(data = CNxTemp_deriv_samples%>%
                    mutate(Mass_per_loss = as.numeric(as.character(Mass_per_loss)),
                           Temperature_C_avg = fct_recode(Temperature_C_avg, "7.5 C" = "7.5", "10 C" = "10","15 C" = "15", "20 C" = "20"))%>%
                    filter(Mass_per_loss>0))+
         geom_line(aes(x= Mass_per_loss, y = exp(md_posterior)), size = 1)+
  geom_line(aes(x= Mass_per_loss, y = exp(lower_posterior)), linetype = 2)+
  geom_line(aes(x= Mass_per_loss, y = exp(upper_posterior)), linetype = 2)+
         facet_grid(initial_CN~Temperature_C_avg, scales = "free")+
  scale_y_continuous(trans="log")+
  geom_hline(yintercept = 1)+
  ylab("Rate of Change (Log C:N/Mass loss)")+
  xlab("Mass Loss (%)")+
  theme_bw()+deriv_plot_theme+
  NULL
CNxTemp_deriv.p
CNxTemp_deriv.p2
```


```{r cowplot N mass and deriv plots}

CN_temp_grid <- plot_grid(CN_T.p,CNxTemp_deriv.p, nrow = 2, labels = "AUTO", label_size = 15)
ggsave2("C:/Users/caleb/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Figures/Exploratory Figures/CNxTemp_r1.pdf", CN_temp_grid, width = 7.8, height = 9.1, dpi = 600)

```

#TN exploration

```{r CN X TN}

CN_tn_ml <- gam(data = det_loc_cnp, CN_ratio ~  
                  ti(Mass_per_loss, m = 2, k = 8) + ti(initial_CN, m = 2)+
                  ti(Mass_per_loss, initial_CN, m = 2)+
               ti(Mass_per_loss,initial_CN, TN_approx, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+ s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "log"),select = TRUE) 

CN_tn <- gam(data = det_loc_cnp, CN_ratio ~  
                  ti(Mass_per_loss, m = 2, k = 8) + ti(initial_CN, m = 2)+
                  ti(Mass_per_loss, initial_CN, m = 2)+
               ti(Mass_per_loss,initial_CN, TN_approx, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+ s(series_index, bs = "re"), 
                   method = "REML", family = Gamma(link = "log"),select = TRUE) 

summary(CN_tn_ml)
appraise(CN_tn)

###saveRDS(CN_tn,"Model_Objects/CN_tn.rds")
##CN_tn <- readRDS("Model_Objects/CN_tn.rds")

CN_TN.pred<-predict_gam(type = "link", CN_tn, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL,Remain_Mass_Category = NULL, initial_CN = c(140,100,70,50,40,25),TN_approx = c(300,600,1000,1500)))%>%
   mutate(TN_approx = as_factor(TN_approx), initial_CN = as_factor(initial_CN))
  


 CN_TN.p<- ggplot(data = CN_TN.pred%>%
                    mutate(TN_approx = fct_recode(TN_approx, "300 ug/L" = "300", "600 ug/L" = "600", "1000 ug/L" = "1000", "1500 ug/L" = "1500")))+
    geom_smooth(se = FALSE, aes(x = Mass_per_loss, y = exp(fit), color = TN_approx, linetype = TN_approx))+
   facet_grid(initial_CN~TN_approx, scales = "free")+
   geom_ribbon(aes(x = Mass_per_loss, ymin = exp(fit - 2*se.fit), ymax = exp(fit + 2*se.fit)), alpha = 0.3)+
   #scale_y_continuous(trans = "log")+
   ylab("C:N (molar)")+
   xlab("Mass loss (%)")+ggtitle("Total Nitrogen x C:N")+theme(legend.position = "none")+
    NULL
 CN_TN.p
 

```

TN significant change in curves - now plot the derivatives. 

This takes a long time run fitted_samples() with 5000 draws (maybe 15 minutes or more), so the object important in running the plot has been saved as an .rds file.

This can be read in with CNxTN_deriv_samples <- readRDS("Model_Objects/CN_tn_derivs.rds") where file connection is set to correct working directory where .rds file is stored.
```{r TN covar curve derivatives}

CNxTN_mass_grid <- expand_grid(Mass_per_loss = seq(0,80, by = 0.2), initial_CN = c(25,40,50,70,100,140), TN_approx = c(300,600,1000,1500), series_index = 578)
n_draws <- 5000
CNxTN_mass_grid_exp <- map_dfr(seq_len(n_draws), ~CNxTN_mass_grid)


CNxTN_gam_fitted <- fitted_samples(CN_tn, exclude = "s(series_index,Mass_per_loss", n = n_draws, newdata = CNxTN_mass_grid, seed = 1)%>%
  bind_cols(CNxTN_mass_grid_exp)%>%mutate(initial_CN = as_factor(initial_CN), TN_approx = as_factor(TN_approx), draw = as_factor(draw), fitted = log(fitted))
#log transformed fitted values because *ratios*


CNxTN_deriv_samples <- CNxTN_gam_fitted%>%arrange(initial_CN, TN_approx)%>%group_by(draw, initial_CN,TN_approx)%>%mutate(fit_dif = (fitted-lag(fitted))/0.2)%>%mutate(Mass_per_loss = as_factor(Mass_per_loss))%>%ungroup()%>%
  group_by(initial_CN,TN_approx, Mass_per_loss)%>%summarize(md_posterior = median(fit_dif), upper_posterior = quantile(fit_dif,0.975,na.rm = TRUE), lower_posterior = quantile(fit_dif,0.025,na.rm = TRUE))

#export object because it took about 15 minutes to create with 5000 draws...
#saveRDS(CNxTN_deriv_samples,"Model_Objects/CN_tn_derivs.rds")

CNxTN_deriv.p<-ggplot(data = CNxTN_deriv_samples%>%
                  mutate(Mass_per_loss = as.numeric(as.character(Mass_per_loss)),
                         TN_approx = fct_recode(TN_approx, "300 ug/L" = "300", "600 ug/L" = "600", "1000 ug/L" = "1000", "1500 ug/L" = "1500"))%>%
                  filter(Mass_per_loss>0))+
         geom_line(aes(x= Mass_per_loss, y = exp(md_posterior)), size = 1)+
  geom_line(aes(x= Mass_per_loss, y = exp(lower_posterior)), linetype = 2)+
  geom_line(aes(x= Mass_per_loss, y = exp(upper_posterior)), linetype = 2)+
         facet_grid(initial_CN~TN_approx, scales = "free")+
  scale_y_continuous(trans="log")+
  geom_hline(yintercept = 1)+
  ylab("Rate of Change (Log C:N/Mass loss)")+
  xlab("Mass Loss (%)")+
  theme_bw()+deriv_plot_theme+
  NULL
CNxTN_deriv.p

```

```{r cowplot N mass and deriv plots}

CN_TN_grid <- plot_grid(CN_TN.p,CNxTN_deriv.p, nrow = 2, labels = "AUTO", label_size = 15)
ggsave2("C:/Users/robbi/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Figures/Exploratory Figures/CNxTN.pdf", CN_TN_grid, width = 7.8, height = 9.1, dpi = 600)

```

#TP exploration
Load model object with CN_tp<-readRDS("Model_Objects/CN_tp.rds")
```{r CN X TP}



CN_tp_ml <- gam(data = det_loc_cnp, CN_ratio ~  
                  ti(Mass_per_loss, m = 2, k = 8) + ti(initial_CN, m = 2)+
                  ti(Mass_per_loss, initial_CN, m = 2)+
               ti(Mass_per_loss,initial_CN, TP_approx, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "ML", family = Gamma(link = "log"),select = TRUE) 


CN_tp <- gam(data = det_loc_cnp, CN_ratio ~  
                  ti(Mass_per_loss, m = 2) + ti(initial_CN, m = 2)+
                  ti(Mass_per_loss, initial_CN, m = 2)+
               ti(Mass_per_loss,initial_CN, TP_approx, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "REML", family = Gamma(link = "log"),select = TRUE) 

summary(CN_tp_ml)
appraise(CN_tp)

###saveRDS(CN_tp,"Model_Objects/CN_tp.rds")

CN_TP.pred<-predict_gam(type = "link", CN_tp, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL,Remain_Mass_Category = NULL, initial_CN = c(140,100,70,50,40,25),TP_approx = c(50,100,300,500)))%>%
   mutate(TP_approx = as_factor(TP_approx), initial_CN = as_factor(initial_CN))
  


 CN_TP.p<- ggplot(data = CN_TP.pred%>%
                  mutate(TP_approx = fct_recode(TP_approx, "50 ug/L" = "50", "100 ug/L" = "100", "300 ug/L" = "300", "500 ug/L" = "500")))+
   geom_smooth(se = FALSE, aes(x = Mass_per_loss, y = exp(fit), color = TP_approx, linetype = TP_approx))+
   facet_grid(initial_CN~TP_approx, scales = "free")+
   geom_ribbon(aes(x = Mass_per_loss, ymin = exp(fit - 2*se.fit), ymax = exp(fit + 2*se.fit)), alpha = 0.3)+
   ylab("C:N (molar)")+
   xlab("Mass loss (%)")+
   ggtitle("Total Phosphorus x C:N")+
   theme(legend.position = "none")+
    NULL
 CN_TP.p
 

```

CNxTP_deriv_samples <- readRDS("Model_Objects/CN_tp_derivs.rds")
```{r TN covar curve derivatives}

CNxTP_mass_grid <- expand_grid(Mass_per_loss = seq(0,80, by = 0.2), initial_CN = c(25,40,50,70,100,140), TP_approx = c(50,100,300,500), series_index = 578)
n_draws <- 5000
CNxTP_mass_grid_exp <- map_dfr(seq_len(n_draws), ~CNxTP_mass_grid)


CNxTP_gam_fitted <- fitted_samples(CN_tp, exclude = "s(series_index,Mass_per_loss", n = n_draws, newdata = CNxTP_mass_grid, seed = 1)%>%
  bind_cols(CNxTP_mass_grid_exp)%>%mutate(initial_CN = as_factor(initial_CN), TP_approx = as_factor(TP_approx), draw = as_factor(draw), fitted = log(fitted))
#log transformed fitted values because *ratios*


CNxTP_deriv_samples <- CNxTP_gam_fitted%>%arrange(initial_CN, TP_approx)%>%group_by(draw, initial_CN,TP_approx)%>%mutate(fit_dif = (fitted-lag(fitted))/0.2)%>%mutate(Mass_per_loss = as_factor(Mass_per_loss))%>%ungroup()%>%
  group_by(initial_CN,TP_approx, Mass_per_loss)%>%summarize(md_posterior = median(fit_dif), upper_posterior = quantile(fit_dif,0.975,na.rm = TRUE), lower_posterior = quantile(fit_dif,0.025,na.rm = TRUE))

#export object because it took about 15 minutes to create with 5000 draws...
#saveRDS(CNxTP_deriv_samples,"Model_Objects/CN_tp_derivs.rds")

CNxTP_deriv.p<-ggplot(data = CNxTP_deriv_samples%>%
                  mutate(Mass_per_loss = as.numeric(as.character(Mass_per_loss)),
                         TP_approx = fct_recode(TP_approx, "50 ug/L" = "50", "100 ug/L" = "100", "300 ug/L" = "300", "500 ug/L" = "500"))%>%
                  filter(Mass_per_loss>0))+
         geom_line(aes(x= Mass_per_loss, y = exp(md_posterior)), size = 1)+
  geom_line(aes(x= Mass_per_loss, y = exp(lower_posterior)), linetype = 2)+
  geom_line(aes(x= Mass_per_loss, y = exp(upper_posterior)), linetype = 2)+
         facet_grid(initial_CN~TP_approx, scales = "free")+
  scale_y_continuous(trans="log")+
  geom_hline(yintercept = 1)+
  ylab("Rate of Change (Log C:N/Mass loss)")+
  xlab("Mass Loss (%)")+
  theme_bw()+deriv_plot_theme+
  NULL
CNxTP_deriv.p

```

```{r cowplot plots}

CN_TP_grid <- plot_grid(CN_TP.p,CNxTP_deriv.p, nrow = 2, labels = "AUTO", label_size = 15)
ggsave2("C:/Users/robbi/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Figures/Exploratory Figures/CNxTP.pdf", CN_TP_grid, width = 7.8, height = 9.1, dpi = 600)

```
#P mass exploratory models 

None of these will fit - tried all covariates with P and C:P. Just not enough data for three-way tensor interactions and REs 

```{r Pi x Temp}

#can't fit
Pi_temp_ml <- gam(data = det_loc, P_mass ~  
                  ti(Mass_per_loss, m = 2) + ti(initial_P, m = 2)+
                  ti(Mass_per_loss, initial_P, m = 2)+
               ti(Mass_per_loss,initial_P, Temperature_C_avg, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs ="re"), 
                   method = "ML", family = Gamma(link = "identity"),select = TRUE) 




```
#CP exploratory models

```{r}
CP_base1 <- gam(data = det_loc_cnp, CP_ratio ~
                        te(Mass_per_loss, initial_CP, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re") + s(series_index, bs = "re"), 
             method = "REML", family = gaussian(link = "log"), select = TRUE)


CP_temp_ml <- gam(data = det_loc_cnp, CP_ratio ~  
                  ti(Mass_per_loss, m = 2) + ti(initial_CP, m = 2)+
                  ti(Mass_per_loss, initial_CP, m = 2)+
               ti(Mass_per_loss,initial_CP, TP_approx, m = 2)+
                        s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs ="re"), 
                   method = "ML", family = gaussian(link = "identity"),select = TRUE) 


```


## Ergosterol exploration


```{r}
ergo_det <- det_loc|>
  filter(Ergosterol_ug.g >0)

erg_m1 <- gam(data = ergo_det, Ergosterol_ug.g ~  
               s(Mass_per_loss, m = 2)+
                s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "REML", family = Gamma(link = "inverse"),select = TRUE) 

erg_m2 <- gam(data = ergo_det, Ergosterol_ug.g ~  
               s(Mass_per_loss, m = 2)+
                s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "REML", family = Gamma(link = "log"),select = TRUE) 

erg_m3 <- gam(data = ergo_det, Ergosterol_ug.g ~  
               s(Mass_per_loss, m = 2)+
                s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "REML", family = Gamma(link = "identity"),select = TRUE) 



map(list(erg_m1,erg_m2,erg_m3),appraise)
map(list(erg_m1,erg_m2,erg_m3),summary)
map(list(erg_m1,erg_m2,erg_m3),AIC)

erg_m1.pred<-predict_gam(type = "link", erg_m1, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL))

erg_m2.pred<-predict_gam(type = "link", erg_m2, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL))

erg_m3.pred<-predict_gam(type = "link", erg_m3, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL))
  
 
 
   erg_m1.p<- ggplot()+
    geom_smooth(data = erg_m1.pred, se = FALSE, aes(x = Mass_per_loss, y = 1/(fit)))+
   geom_ribbon(data = erg_m1.pred, aes(x = Mass_per_loss, ymin = 1/(fit - 2*se.fit), ymax = 1/(fit + 2*se.fit)), alpha = 0.3)+
   geom_point(data = ergo_det, aes(x = Mass_per_loss, y = Ergosterol_ug.g))+
     scale_x_continuous(breaks = c(0,20,40,60,75))+
   ylab(bquote(bold(paste("Ergosterol (",mu, "g/g litter mass)"))))+
   xlab("Mass Loss (%)")+
     theme_bw()+plot_theme+theme(axis.title.x = element_blank())+
    NULL
 erg_m1.p
 
ergo_det_na<-ergo_det|>filter(!is.na(Ergosterol_ug.g))|>mutate(series_index = as.numeric(as.character(series_index)))

unique(ergo_det_na$series_index)|>length() 
```

```{r ergo curve derivative}

ergo_mass_grid <- expand_grid(Mass_per_loss = seq(0,75, by = 0.2), series_index = 578)
n_draws <- 10000
ergo_mass_grid_exp <- map_dfr(seq_len(n_draws), ~ergo_mass_grid)


ergo_gam_fitted <- fitted_samples(erg_m1, exclude = "s(series_index,Mass_per_loss", n = n_draws, newdata = ergo_mass_grid, seed = 1, scale = "response")%>%
  bind_cols(ergo_mass_grid_exp)%>%mutate(draw = as_factor(draw), fitted = fitted)
#log transformed fitted values because *ratios*


ergo_deriv_samples <- ergo_gam_fitted%>%group_by(draw)%>%mutate(fit_dif = (fitted-lag(fitted))/0.2)%>%mutate(Mass_per_loss = as_factor(Mass_per_loss))%>%ungroup()%>%
  group_by(Mass_per_loss)%>%summarize(md_posterior = median(fit_dif), upper_posterior = quantile(fit_dif,0.975,na.rm = TRUE), lower_posterior = quantile(fit_dif,0.025,na.rm = TRUE))

#export object because it took about 15 minutes to create with 5000 draws...
#saveRDS(CNxTP_deriv_samples,"Model_Objects/CN_tp_derivs.rds")

ergo_deriv.p<-ggplot(data = ergo_deriv_samples%>%
                  mutate(Mass_per_loss = as.numeric(as.character(Mass_per_loss)))%>%
                  filter(Mass_per_loss>0))+
         geom_line(aes(x= Mass_per_loss, y = md_posterior), size = 1)+
  geom_line(aes(x= Mass_per_loss, y = lower_posterior), linetype = 2)+
  geom_line(aes(x= Mass_per_loss, y = upper_posterior), linetype = 2)+
  geom_hline(yintercept = 1)+
  scale_x_continuous(breaks = c(0,20,40,60,75))+
  scale_y_continuous(limits = c(-60,55))+
  ylab(expression(atop(bold("Rate of Change"), bold("(Ergosterol/Mass loss)"))))+
  xlab("Mass Loss (%)")+
  theme_bw()+deriv_plot_theme+
  NULL
ergo_deriv.p

```


```{r cowplot ergo plots}
ergo_grid<-plot_grid(erg_m1.p,ergo_deriv.p, nrow = 2, labels = "AUTO", label_size = 20)

ggsave2("C:/Users/caleb/Dropbox/Detrital Nutrients Synth/Manuscript Dev/Temporal Trends ms/Figures/DetNut_Fig4_Ergosterol.pdf", ergo_grid, width = 10, height = 10, dpi = 600)

```

```{r}

ergtemp_m1 <- gam(data = ergo_det, Ergosterol_ug.g ~  
               s(Mass_per_loss, m = 2)+ti(Mass_per_loss,Temperature_C_avg)+
                s(series_index, Mass_per_loss, bs = "re")+s(series_index, bs = "re"), 
                   method = "REML", family = Gamma(link = "inverse"),select = TRUE) 
 
  erg_temp.pred<-predict_gam(type = "link", ergtemp_m1, length_out = 5000,  exclude_terms = c("s(series_index)", "s(series_index, Mass_per_loss)"), values = list(series_index = NULL, Remain_Mass_Category = NULL, Temperature_C_avg = c(10,15,20,25)))%>%mutate(Temperature_C_avg = as_factor(Temperature_C_avg))
  
appraise(ergtemp_m1)
AIC(ergtemp_m1,erg_m1)

 erg_temp.p<- ggplot()+
    geom_smooth(data = erg_temp.pred, se = FALSE, aes(x = Mass_per_loss, y = 1/fit, color = Temperature_C_avg, linetype = Temperature_C_avg))+
   #facet_grid(Temperature_C_avg~., scales = "free")+
   #geom_ribbon(data = erg_temp.pred, aes(x = Mass_per_loss, ymin = 1/(fit - 2*se.fit), ymax = 1/(fit + 2*se.fit)), alpha = 0.3)+
    geom_abline(slope = -0.01, intercept = 1)+
    NULL
 erg_temp.p

 
  erg_temp.p2<- ggplot()+
    geom_smooth(data = erg_temp.pred, se = FALSE, aes(x = Mass_per_loss, y = 1/fit, color = Temperature_C_avg, linetype = Temperature_C_avg))+
   #facet_grid(Temperature_C_avg~., scales = "free")+
   geom_ribbon(data = erg_temp.pred, aes(x = Mass_per_loss, ymin = 1/(fit - 2*se.fit), ymax = 1/(fit + 2*se.fit),color = Temperature_C_avg), alpha = 0.3)+
    geom_abline(slope = -0.01, intercept = 1)+
    scale_y_continuous(limits = c(0,1000))+
    NULL
 erg_temp.p2
 

```




```{r}
dupe <- det_loc|>
  group_by(First_Author, Publication_Title, Journal, series_index)%>%
  mutate(series_length = n())|>relocate(series_length)|>
  nest()%>%
  mutate(System = map_chr(data, ~first(.$System)),
         Lotic_Lentic = map_chr(data, ~first(.$Lotic_Lentic)),
         Detritus_Condition = map_chr(data, ~first(.$Detritus_Condition)),
         Detritus_Species = map_chr(data, ~first(.$Detritus_Species)),
         Detritus_Type = map_chr(data, ~first(.$Detritus_Type)),
         Decay_Method = map_chr(data, ~first(.$Decay_Method)),
         initial_C = map_dbl(data, ~first(.$initial_C)),
         initial_N = map_dbl(data, ~first(.$initial_N)),
         initial_P = map_dbl(data, ~first(.$initial_P)),
         initial_CN = map_dbl(data, ~first(.$initial_CN)),
         initial_CP = map_dbl(data, ~first(.$initial_CP)),
         initial_NP = map_dbl(data, ~first(.$initial_NP)),
         series_length = map_dbl(data, ~first(.$series_length)),
        CNP_Ratio_Type = map_chr(data, ~first(.$CNP_Ratio_Type)))

dupe_p <- filter(dupe, !is.na(initial_P))|>ungroup()|>summarize(points = sum(series_length)) #99 series, 484
dupe_n <- filter(dupe, !is.na(initial_N))|>ungroup()|>summarize(points = sum(series_length)) #119 series, 523
dupe_cp <- filter(dupe, !is.na(initial_CP)&CNP_Ratio_Type=="molar")|>ungroup()|>summarize(points = sum(series_length)) #91, 357 points
dupe_cn <- filter(dupe, !is.na(initial_CN))|>ungroup()|>summarize(points = sum(series_length)) #141 series, 601 points

dupe2 <- filter(det_loc, !is.na(N_mass))

```
